Semantic browsing of sound databases \\ without keywords

### 2016-06-01 responseToReviewers.tex

\section{Responses to Reviewer 1}

\begin{enumerate}

\item \emph{[Are] the sounds heard counter was reset for each search? That is, whether the total and unique heard sounds are related to the whole experiment or an average value per target?}

The total and unique heard sounds are related to the whole experiment.

@@

\item \emph{Was this controlled by the study or was the study simply time limited?}

Heu de quoi ?

\item \emph{Perhaps in the post-hoc analyses it would be useful to know if FSD produces significant results in pair-wise tests with Bonferroni correction (since this is, albeit simple, still a multiple comparison test).}

Post hoc analyses are now down using the Bonferroni procedure. All the p-values are updated. Considering the urban dataset, we don't achieve statistical significance for the search durations any-more. But this does not change the paper conclusions.

### 2016-06-01 responseToReviewers.tex

\section{Responses to Reviewer 1}

\begin{enumerate}

\item \emph{[Are] the sounds heard counter was reset for each search? That is, whether the total and unique heard sounds are related to the whole experiment or an average value per target?}

The total and unique heard sounds are related to the whole experiment. The Section 5.2 of the updated manuscript is now more explicit in this respect.

\item \emph{How many target sounds were assessed by each subject? Was this controlled by the study or was the study simply time limited?}

All the subjects had to assess 13 target sounds, thus controlled by the study, see Section 4.2 of the updated manuscript.

### 2016-06-07 lafaySsf2016.tex
polishing response
include nicolas review, merge Fig 5 and 7, increase font size

\usepackage{color,url}
\newcommand{\gl}[1]{\textcolor{red}{Gregoire : #1}}
\newcommand{\mr}[1]{\textcolor{green}{Mathias : #1}}
\newcommand{\ml}[2][]{\textcolor{blue}{#1 #2}}

%\renewcommand{\baselinestretch}{1.5}

@@

%Author Info.
\authorgroup{
\author{Gr\'egoire Lafay$^1$}, \author{Nicolas Misdariis$^2$},
\author{Mathieu Lagrange$^1$}, \author{Mathias Rossignol$^2$}
\email{(gregoire.lafay@irccyn.ec-nantes.fr)}
\affil{IRCCyN, Ecole Centrale de Nantes $^1$ \quad\quad\quad\quad IRCAM, STMS, UPMC $^2$}
}

%\author{
%\alignauthor
% Gregoire Lafay\\
%       \affaddr{IRCCyN, Ecole Centrale de Nantes}\\
%       \affaddr{1 rue de la Noe}\\
%       \affaddr{Nantes, France} 
%       %\email{gregoire.lafay@irccyn.ec-nantes.fr} 
%\alignauthor 
% Nicolas Misdarris\\
%       \affaddr{IRCAM}\\
%       \affaddr{1 Place Igor\-Stravinsky}\\
%       \affaddr{Paris, France} \\
%       \email{}
%\and
%\alignauthor 
%Mathieu Lagrange\\
%       \affaddr{IRCCyN, Ecole Centrale de Nantes}\\
%       \affaddr{1 rue de la Noe}\\
%       \affaddr{Nantes, France} \\
%\alignauthor 
% Mathias Rossignol\\
%       \affaddr{IRCAM}\\
%       \affaddr{1 Place Igor\-Stravinsky}\\
%       \affaddr{Paris, France} \\
%       \email{}
%}

@@

We present and study the efficiency of two sound presentation strategies that organize sounds spatially so as to reflect an underlying semantic hierarchy. For the sake of comparison, we also consider a display whose spatial organization is based on acoustic cues. Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first being composed of sounds of urban environment and the second of sounds produced by musical instruments. Coherent results achieved by considering the two different corpora demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of efficiency of learning. 
\end{abstract}

@@

With the growing capability of recording and storage devices, the problem of indexing large databases of audio has recently been the object of much attention \cite{Wold1996}. Most of that effort is dedicated to the automatic inference of indexing metadata from the actual audio recording \cite{Zhang1999, tzanetakis2002musical}; in contrast, the ability to browse such databases in an effective manner has been less considered.

@@

\begin{enumerate}
\item Sounds, as many others things, can be described in many ways. Sound may be designated by their
sources (a car door), as well as by the action of those sources (the slamming of a car door) or their environments (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}. Designing an effective keyword-based search system requires an accurate description of each sound, which has to be tunable to the sound representation of each user.
\item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their browsing and final selection.
\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound, such as its ambiance, is notoriously ambiguous and subject to cultural specificities.
\item Unless considerable time and resources are invested into developing a multilingual interface, any system based on verbal descriptions can only be used with reduced performance by non-native speakers of the chosen language.
\end{enumerate}
To circumvent those issues, not relying on keywords is desirable. However, conveying semantics remains a necessity for ease of browsing, which raises the question of what alternate means, other than written language, can be used to that end.

@@

Their effectiveness is studied through a search-based task whose aim is to find a given target sound by browsing the database using the display under evaluation. Two corpora are considered for this study. The first one is composed of sounds recorded in a urban environment and the second one is composed of sounds of musical instruments.

@@

Similarly, the latter also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \textit{free-aerophone} $>$ \textit{brassy} $>$ \textit{trumpet}).

@@

%\begin{figure}[t]
%\begin{center}
%\includegraphics[scale=0.18]{gfx/xp3.png} 
%\end{center}
%\caption{\label{figXP3} Acoustical Display (AD).}
%\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.5\textwidth]{gfx/urban_mds.png} 
\end{center}
\caption{\label{figXP3urban} Acoustical Display for the urban sound dataset (AD). Colors and legend are removed during the experiment.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.5\textwidth]{gfx/music_mds.png} 
\end{center}
\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment.}
\end{figure}

Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD)%, shown in Figures~\ref{figXP2} and~\ref{figXP1}, respectively. 
Both display  consider the hierarchical structure of the dataset to organize sounds. Each sound class is represented by a filled circle. Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown in Figure~\ref{figSSF}. Thus, subclasses belonging to the same class are close to each others. Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space. Depending on the display, the user can either access each leaf class directly, or has to click through intermediate levels of the hierarchy.

@@

%\begin{figure}[t!]
%\begin{center}
%\includegraphics[scale=0.30]{gfx/XP2clean.pdf} 
%\end{center}
%\caption{\label{figXP2} Full Semantic Display (FSD): implicit hierarchical organization of semantic classes.}
%\end{figure}
%
%\begin{figure*}[t]
%\begin{center}
%\includegraphics[scale=0.4]{gfx/XP1clean2.pdf} 
%\end{center}
%\caption{\label{figXP1} Progressive Semantic Display (PSD):  explicit hierarchical organization of semantic classes. (a) initial folded version; (b) partly folded version; (c) unfolded version.}
%\end{figure*}

@@

During this experiment, the three displays are compared using two corpora: the urban sound corpus and the musicla sound corpus. By comparing AD and xSD (the two semantic displays), the goal is to evaluate the  relative efficiencies of semantic based and acoustic based spatial configurations. By comparing FSD and PSD, we study the impact of an enforced hierarchical exploration of the corpus by the user. The goal is to check if constraining the user to first browse the high levels of the hierarchy helps him to grasp and memorize the spatial configuration and the organization of the sound classes.

@@

\subsection{Data Collection}
Four types of data are collected during the experiment:
\begin{itemize}
\item the total duration of the entire experiment. It includes breaks between two target sound searches and it is called the \textit{absolute duration}.
\item the duration of each search. The sum of the 13 search durations, which is the \textit{absolute duration} minus the break times between two target sound searches, is called the \textit{total search duration}.
\item the name of each sound which has been heard.
\item the time at which each sound has been heard.  
\end{itemize}

@@

A crowdsourcing approach is adopted. The experiment is designed to be performed using the \textit{Chrome} web browser. The link to the experiment has been sent to three  mailing lists, namely \textit{Music-Ir, Auditory} and an internal IRCAM mailing list. Subjects are only allowed to perform the experiment once, and on one interface only. Data were automatically collected server-side at the end of the experiments. Subjects were asked to use headphones. All sounds of the dataset are normalized to the same root mean square (RMS) level.

@@

60 subjects have completed the experiment for the urban sound dataset, 20 for each interface. 43 subjects have completed the experiment for the urban sound dataset: 15 for the PSD interface, 13 for the FSD interface and 15 for the AD interface.

\ml{to update before final submission}

@@

We use one-way ANOVA to test for statistical significance. Post hoc analyses are done using two samples \textit{t} tests with Bonferroni correction. All statistical analyses are performed at the 5\% significance level.% using native functions of the MATLAB computing environment.

@@

%\begin{table}[t]
%\caption{\label{tab1} Criterion used to detect outliers.}
%\begin{center}
%\begin{tabular}{ccccc}
%criterion      &  1  &  2  &  3  &  4  \\
%\hline
%outlier 1 (PD) & --- & --- & --- &  x  \\
%outlier 2 (PD) &  x  & --- & --- & --- \\  
%outlier 3 (AD) & --- & --- & --- &  x \\
%outlier 4 (AD) &  x  &  x  &  x  & --- \\
%outlier 5 (AD) &  x  & --- & --- & --- \\
%\hline
%\end{tabular}
%\end{center}
%\end{table}

Using the $IQR$ method, 5 subjects are detected as outliers for the urban experiment, and removed from the analysis. 2 of those subjects used PSD and 3 subjects AD. %Table~\ref{tab1} shows the criteria used to detect the 5 outliers. 
3 subjects are detected by considering the \textit{absolute duration}. They spent respectively 50 minutes, 5 hours and 17.5 hours to complete the experiment. 1 subject is detected by considering the \textit{total search duration} (46 minutes), and an other by observing the \textit{total number of heard sounds} (1800 heard sounds, roughly 12 times the total size of the corpus). And lastly 2 subjects are detected by observing the maximum number of times they had to hear a target sound before finding it (4 and 8 times).

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.4]{gfx/analyse_urban1.eps} 
\end{center}
\caption{\label{fig1urban} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.4]{gfx/analyse_urban2.eps} 
\end{center}
\caption{\label{fig2urban} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search. Lines are linear regression fits.}
\end{figure}

@@

\begin{itemize}
\item the \textit{total search duration}
\item the number of heard sounds
\item the number of unique sounds heard. By "unique" we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
\end{itemize}

@@

The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1urban}. The type of display has a significant effect on the the \textit{total search duration} ($F[2,52]=3.63; p<0.05$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of unique heard sounds ($F[2,52]=8.65; p<0.01$).

Post hoc analysis on the total search time (Figure~\ref{fig1urban}.a) shows no significant differences between the three interfaces (PSD-FSD: $p=0.09$; PSD-AD: $p=1$; FSD-AD: $p=0.07$), although the FSD interface achieves the best results.

Concerning the numbers of heard sounds (Figure~\ref{fig1urban}.b), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$),  whereas PSD and FSD as well as PSD and AD show similar outcomes (PSD-FSD: $p=0.06$; PSD-AD: $p=1$).

Figure~\ref{fig1urban}.c shows the results for the number of unique heard sounds. This time the results of AD are significantly lower than those of both PSD and FSD (PSD-AD: $p<0.01$; FSD-AD: $p<0.01$).  For AD, 75\% of the subjects heard more than 140 sounds, and 25\% heard at least 148 sounds, that is almost the entire database. Considering PSD, 75\% of subjects heard less than 134 sounds, versus 143 for FSD. This time PSD and FSD perform equivalently (PD-FD: $p=1$).

@@

Figure~\ref{fig2urban}.a shows the evolution of the average duration of each target sound search observed for AD, FSD and PSD. As shown by a linear regression of the data, there is an overall increase of efficiency with respect to the index of the search. For the duration and the number of heard sounds (Figure~\ref{fig2urban}.b), the AD and PSD perform equivalently, whereas the FSD exhibit better performance. In the case of the number of unique heard sounds (Figure~\ref{fig2urban}.c), the two semantic displays are equivalent. This latter result suggest that for both semantic displays, users have parsed the same range of the dataset, although users of FSD have done so by listening to fewer sounds and more quickly.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.4]{gfx/analyse_music1.eps} 
\end{center}
\caption{\label{fig1Music} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.4]{gfx/analyse_music2.eps} 
\end{center}
\caption{\label{fig2Music} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search. Lines are linear regression fits.}
\end{figure}

%\gl{merge fig. 5 and 7}
% if space is need fig 1 could be merged

@@

The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1Music}. The type of display has a significant effect on the the \textit{total search duration} ($F[2,38]=10.25; p<0.01$), the number of heard sounds ($F[2,38]=13.15; p<0.01$) and the number of unique heard sounds ($F[2,38]=6.39; p<0.01$).

Considering the \textit{total search duration} (Figure~\ref{fig1Music}.a), post hoc analyses shows that PSD and FSD perform better than the AD interfaces (PSD-AD: $p<0.05$; FSD-AD: $p<0.01$;). No significant difference is found between PSD and FSD (PSD-FSD: $p=0.32$).

Figure~\ref{fig1Music}.b shows the results for the numbers of heard sounds. This time AD and PSD exhibit similar results (PSD-AD: $p<0.43$), while being both outperformed by FSD (FSD-PSD: $p<0.01$; FSD-AD: $p<0.01$).

Considering the numbers of unique heard sounds (Figure~\ref{fig1Music}.c), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$), while both PSD \emph{vs.} FSD and PSD \emph{vs.} AD show similar outcomes (PSD-FSD: $p=0.6$; PSD-AD: $p=0.9$). Those results confirms the trends observed for the urban sounds (see Section~\ref{resultsu}):

@@

Considering that the musical corpus have perhaps a obvious semantic organization, PSD and FSD significantly reduces the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset.

@@

>>>>

\usepackage{color,url}
\newcommand{\gl}[1]{\textcolor{red}{gl: #1}}
\newcommand{\mr}[1]{\textcolor{green}{Mathias : #1}}
\newcommand{\ml}[2][]{\textcolor{blue}{#1 #2}}
\newcommand{\nm}[1]{\textcolor{magenta}{nm: #1}}
%\renewcommand{\baselinestretch}{1.5}

@@

%Author Info.
\authorgroup{
\author{Gr\'egoire Lafay$^1$}, \author{Nicolas Misdariis$^2$},
\author{Mathieu Lagrange$^1$}, \author{Mathias Rossignol$^2$}
\email{(gregoire.lafay@irccyn.ec-nantes.fr)}
\affil{IRCCyN, Ecole Centrale de Nantes $^1$ \quad\quad\quad\quad STMS Ircam-CNRS-UPMC $^2$}
}

@@

We present and study the efficiency of two sound presentation strategies that organize sounds spatially so as to reflect an underlying semantic hierarchy. For the sake of comparison, we also consider a display whose spatial organization is based on acoustic cues. Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first being composed of sounds of urban environment \nm{je mettrais plutot: "environmental sounds from urban"} and the second of sounds produced by musical instruments. Coherent results achieved by considering the two different corpora demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of \gl{learning efficiency}. 
\end{abstract}

@@

With the growing capability of recording and storage devices, the problem of indexing large databases of audio \nm{"audio databases" ?} has recently been the object of much attention \cite{Wold1996}. Most of that effort is dedicated to the automatic inference of indexing metadata from the actual audio recording \cite{Zhang1999, tzanetakis2002musical}; in contrast, the ability to browse such databases in an effective manner has been less considered.

@@

\begin{enumerate}
\item Sounds, as many others things, can be described in many ways. \nm{ceci est valable pour des sons de nature "causale", i.e. dont la source est identifiable; pour les sons abstraits, la semantique de description est davanage de l'ordre de l'acoustique (briallnt, aigu, sourd, rond, et.) voir p. ex. Lemaitre, G., Houix, O., Misdariis, N., Susini, P., « Listener expertise and sound identification influence the categorization of environmental sounds », Journal of Experimental Psychology: Applied, 2010, vol. 16, n° 1, pp. 16-32} Sound may be designated by their
sources (a car door), as well as by the action of those sources (the slamming of a car door) or their environments (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}. Designing an effective keyword-based search system requires an accurate description of each sound, which has to be tunable to the sound representation of each user.
\item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their browsing and final selection.
\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound, such as its ambiance, is notoriously ambiguous and subject to cultural specificities.
\item Unless considerable time and resources are invested into developing a multilingual interface, any system based on verbal descriptions can only be used with reduced performance by non-native speakers of the chosen language.
\end{enumerate}
To circumvent those issues, not relying on keywords is desirable. However, conveying semantics remains a necessity for ease of browsing, which raises the question of what alternate means, other than written language, can be used to that end.

@@

Their effectiveness is studied through a search-based task whose aim is to find a given target sound by browsing the database using the display under evaluation. Two corpora are considered for this study. The first one is composed of sounds recorded in a urban environment and the second one is composed of sounds of musical \nm{"produced by musical ..." ?} instruments.

@@

Similarly, the latter \nm{"the latter corpus" / "the second corpus" ?} also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \gl{\textit{brass} $>$ \textit{ordinario} $>$ \textit{trumpet}}).

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/urban_mds.png} 
\end{center}
\caption{\label{figXP3urban} Acoustical Display for the urban sound dataset (AD). Colors and legend are removed during the experiment.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/music_mds.png} 
\end{center}
\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment. \gl{remove ?}}
\end{figure}

Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD). Both display  consider the hierarchical structure of the dataset to organize sounds. Each sound class is represented by a filled circle. Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown in Figure~\ref{figSSF}. Thus, subclasses belonging to the same class are close to each others. Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space. Depending on the display, the user can either access each leaf class directly, or has to click through intermediate levels of the hierarchy.

@@

@@

During this experiment, the three displays are compared using two corpora: the urban sound corpus and the musical sound corpus. By comparing AD and xSD (the two semantic displays), the goal is to evaluate the  relative efficiencies of semantic based and acoustic based spatial configurations. By comparing FSD and PSD, we study the impact of an enforced hierarchical exploration of the corpus by the user. The goal is to check if constraining the user to first browse the high levels of the hierarchy helps him to grasp and memorize the spatial configuration and the organization of the sound classes.

@@

\subsection{Data Collection}
Four types of data are collected during the experiment:
\begin{itemize}
\item the total duration of the entire experiment. It includes breaks between two target sound searches and it is called the \textit{absolute duration}.
\item the duration of each search. The sum of the 13 search durations, which is the \textit{absolute duration} minus the break times between two target sound searches, is called the \textit{total search duration}.
\item the name of each sound which has been heard. \nm{... and so the number of heard sound (to make the link with sect. 4.6 below)}
\item the time at which each sound has been heard.  
\end{itemize}

@@

A crowdsourcing approach is adopted. The experiment is designed to be performed using the \textit{Chrome} web browser. The link to the experiment has been sent to three  mailing lists, namely \textit{Music-Ir, Auditory} and an internal IRCAM mailing list. Subjects are only allowed to perform the experiment once, and on one interface only. \nm{may be worth specifying how this is done : IP id tracking ?} Data were automatically collected server-side at the end of the experiments. Subjects were asked to use headphones. All sounds of the dataset are normalized to the same root mean square (RMS) level.

@@

60 subjects have completed the experiment for the urban sound dataset, 20 for each interface. \gl{48 subjects have completed the experiment for the urban sound dataset: 17 for the PSD interface, 16 for the FSD interface and 15 for the AD interface.}

@@

We use one-way ANOVA to test for statistical significance. Post hoc analyses are done using two samples \textit{t} tests with Bonferroni correction. All statistical analyses are performed at the 5\% significance level.

@@

Using the $IQR$ method, 5 subjects \nm{sauf erreur, le total des sujets outliers fait 7 (3+1+1+2) alors qu'il est annonce 5 en debut de paragraphe} \gl{non il y a bien 5 outliers} are detected as outliers for the urban experiment, and removed from the analysis. 2 of those subjects used PSD and 3 subjects AD. 3 subjects are detected by considering the \textit{absolute duration}. They spent respectively 50 minutes, 5 hours and 17.5 hours to complete the experiment. 1 subject is detected by considering the \textit{total search duration} (46 minutes), and another by observing the \textit{total number of heard sounds} (1800 heard sounds, roughly 12 times the total size of the corpus). And lastly 2 subjects are detected by observing the maximum number of times they had to hear a target sound before finding it (4 and 8 times).

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/urbanMusic1.eps} 
\end{center}
\caption{\label{urbanMusic} \gl{Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering the urban dataset (a,b,c) or the music dataset (d,e,f)}}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/analyse_urban2.eps} 
\end{center}
\caption{\label{fig2urban} \gl{evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the urban dataset. Lines are linear regression fits.}}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/analyse_music2.eps} 
\end{center}
\caption{\label{fig2Music} \gl{evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the music dataset. Lines are linear regression fits.}}
\end{figure}

@@

\begin{itemize}
\item the \textit{total search duration}
\item the number of heard sounds during the whole experiment
\item the number of unique sounds heard during the whole experiment. By "unique" we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
\end{itemize}

@@

The averages and standard deviations for the three metrics are presented in Figure~\ref{urbanMusic}. The type of display has a significant effect on the the \textit{total search duration} ($F[2,52]=4; p<0.05$), the number of heard sounds ($F[2,52]=6; p<0.01$) and the number of unique heard sounds ($F[2,52]=9; p<0.01$).

Post hoc analysis on the total search duration (Figure~\ref{urbanMusic}.a) shows no significant differences between the three interfaces (PSD-FSD: $p=0.09$; PSD-AD: $p=1$; FSD-AD: $p=0.07$), although the FSD interface achieves the best results.

Concerning the numbers of heard sounds (Figure~\ref{urbanMusic}.b), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$),  whereas PSD and FSD as well as PSD and AD show similar outcomes (PSD-FSD: $p=0.06$; PSD-AD: $p=1$).

Figure~\ref{urbanMusic}.c shows the results for the number of unique heard sounds. This time the results of AD are significantly lower \nm{higher ?} than those of both PSD and FSD (PSD-AD: $p<0.01$; FSD-AD: $p<0.01$).  For AD, 75\% of the subjects heard more than 140 sounds, and 25\% heard at least 148 sounds, that is almost the entire database. Considering PSD, 75\% of subjects heard less than 134 sounds, versus 143 for FSD. This time PSD and FSD perform equivalently (PSD-FSD: $p=1$).

@@

Figure~\ref{fig2urban}.a shows the evolution of the average duration of each target sound search observed for AD, FSD and PSD. As shown by a linear regression of the data, there is an overall increase of efficiency with respect to the index of the search. For the duration and the number \nm{"the number" ?} of heard sounds (Figure~\ref{fig2urban}.b), the AD and PSD perform equivalently \nm{mais pt-etre faudrait-il mentionner ici que la variation autour de la droite de regression de PSD est beaucoup plus importante ... ?}, whereas the FSD exhibit better performance. In the case of the number of unique heard sounds (Figure~\ref{fig2urban}.c), the two semantic displays are equivalent \nm{idem}. This latter result suggests that for both semantic displays, users have parsed the same range of the dataset, although users of FSD have done so by listening to fewer sounds and more quickly.

@@

@@

The averages and standard deviations for the three metrics are presented in Figure~\ref{urbanMusic}. The type of display has a significant effect on the the \textit{total search duration} ($F[2,43]=13; p<0.01$), the number of heard sounds ($F[2,43]=14; p<0.01$) and the number of unique heard sounds ($F[2,43]=5; p<0.05$).

Considering the \textit{total search duration} (Figure~\ref{urbanMusic}.d), post hoc analyses shows that PSD and FSD perform better than the AD interfaces (PSD-AD: $p<0.01$; FSD-AD: $p<0.01$;). No significant difference is found between PSD and FSD (PSD-FSD: $p=0.20$).

Figure~\ref{urbanMusic}.e shows the results for the numbers of heard sounds. This time AD and PSD exhibit similar results (PSD-AD: $p=0.19$), while being both outperformed by FSD (FSD-PSD: $p<0.01$; FSD-AD: $p<0.01$).

Considering the numbers of unique heard sounds (Figure~\ref{urbanMusic}.f), FSD significantly outperforms the AD interface (FSD-AD: $p<0.05$), while both PSD \emph{vs.} FSD and PSD \emph{vs.} AD show similar outcomes (PSD-FSD: $p=0.06$; PSD-AD: $p=1$). Those results confirms the trends observed for the urban sounds (see Section~\ref{resultsu}):

@@

Considering that the musical corpus have perhaps an obvious semantic organization, PSD and FSD significantly reduces the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset.

@@

\nm{Cela manque pt-etre d'une partie Discussion / Synthese ? Et la conclusion manque pt-etre un peu de perspectives : vers simScene, par exemple ?}
